The Impact of News Sentiment on MSFT

### 2023-12-18 MSFT_with_vadar_analysis/MSFT_vadar_analysis.Rmd
VAR with vadar sentiment analysis

# Analysis:
First, we import the data and convert it to a time series so that we may plot it.
After importing and converting our dataset into a workable format, we observed noticeable noise in the average_polarity data. To address this, we applied a LOESS smoothing technique. This method allows us to smooth out short-term fluctuations and highlight longer-term trends in sentiment data, crucial for correlating with stock price movements. The choice of the smoothing parameter (span) was iteratively refined to balance detail and smoothness

```{r plot_chunk, fig.height=10, fig.width=14, echo=FALSE, include=TRUE}
MSFT <- read.csv("data/final_data.csv")
MSFT$Index <- as.Date(MSFT$Index)
MSFT$MSFT.Adjusted <- as.numeric(as.character(MSFT$MSFT.Adjusted))
MSFT$average_polarity <- as.numeric(as.character(MSFT$average_polarity))
MSFT_ts <- zoo(MSFT$MSFT.Adjusted, order.by = MSFT$Index)

@@

ix.na = is.na(MSFT$average_polarity)
problematic_values <- MSFT$average_polarity[ix.na]
MSFT$average_polarity[problematic_values] <- NA
# Replace with 0
MSFT$average_polarity[problematic_values] <- 0
MSFT$Index_numeric <- as.numeric(MSFT$Index - as.Date("2000-01-01"))
MSFT$smoothed_polarity_loess <- with(MSFT, loess(average_polarity ~ Index_numeric, span = 0.2)$fitted)

@@

  geom_line(aes(y = smoothed_polarity_loess * 2000, color = "Sentiment Average Score"), size = 1) +
  ylab("Sentiment Average Score") +

@@

```{r echo=FALSE, include=TRUE}
news_ts <- zoo(MSFT$average_polarity, order.by = MSFT$Index)

@@

MSFT_VAR <- VAR(MSFT_VAR_Data, ic="AIC", lag.max = 24)
coeftest(MSFT_VAR)
```
We see that almost none of the coefficients have a statistical significance as their p-values are all above 0.05. An exception is the second lag of MSFT returns. It seems to have a slight negative effect on the current price of MSFT. However, there is only small evidence.

@@

```{r echo=FALSE, include=TRUE}
causality(MSFT_VAR, cause="News_Sentiment")["Granger"] 
```
The p-value of 0.34 confirms that the gathered News sentiments are not a causal driver for the MSFT stock price. This leads us to reject our hypothesis.

>>>>

# Analysis:
First, we import the data and convert it to a time series so that we may plot it.
After importing and converting our dataset into a workable format, we observed noticeable noise in the sentiment_average_score data. To address this, we applied a LOESS smoothing technique. This method allows us to smooth out short-term fluctuations and highlight longer-term trends in sentiment data, crucial for correlating with stock price movements. The choice of the smoothing parameter (span) was iteratively refined to balance detail and smoothness

```{r plot_chunk, fig.height=10, fig.width=14, echo=FALSE, include=TRUE}
MSFT <- read.csv("final_data_vadar.csv")
MSFT$Index <- as.Date(MSFT$Index)
MSFT$MSFT.Adjusted <- as.numeric(as.character(MSFT$MSFT.Adjusted))
MSFT$sentiment_average_score <- as.numeric(as.character(MSFT$sentiment_average_score))
MSFT_ts <- zoo(MSFT$MSFT.Adjusted, order.by = MSFT$Index)

@@

ix.na = is.na(MSFT$sentiment_average_score)
problematic_values <- MSFT$sentiment_average_score[ix.na]
MSFT$sentiment_average_score[problematic_values] <- NA
# Replace with 0
MSFT$sentiment_average_score[problematic_values] <- 0
MSFT$Index_numeric <- as.numeric(MSFT$Index - as.Date("2000-01-01"))
MSFT$smoothed_polarity_loess <- with(MSFT, loess(sentiment_average_score ~ Index_numeric, span = 0.2)$fitted)

@@

  geom_line(aes(y = sentiment_average_score, color = "Sentiment Average Score"), size = 1) +
  ylab("Sentiment Average Score") +

@@

```{r echo=FALSE, include=TRUE}
news_ts <- zoo(MSFT$sentiment_average_score, order.by = MSFT$Index)

@@

MSFT_VAR <- VAR(MSFT_VAR_Data, ic="AIC", lag.max = 24)
coeftest(MSFT_VAR)
```
The coefficient test reveals some interesting insights into the dynamics between MSFT stock returns and news sentiment. In particular, several lagged values of news sentiment exhibit statistical significance, suggesting a nuanced, albeit modest, effect on stock returns. For example, both the second and fifth lags of sentiment show a negative impact, while the eighth lag shows a more pronounced negative impact. This pattern is mirrored by the eleventh and fifteenth lags, but with less statistical strength. Conversely, the impact of news sentiment is not uniform across all lags, as some show a positive relationship, although these are not statistically significant at conventional levels.

These findings suggest that while news sentiment does not consistently drive stock returns, it does have some predictive power at certain intervals. This time effect could be due to the market's lagged reaction to news sentiment or to the complex interplay between various market forces and investor psychology.

@@

```{r echo=FALSE, include=TRUE}
causality(MSFT_VAR, cause="News_Sentiment")["Granger"] 
```
he Granger causality test provided a new perspective on the relationship between news sentiment and MSFT stock returns. Contrary to our initial hypothesis, the test indicates that news sentiment may indeed have a Granger causal effect on stock returns, as indicated by a p-value of 0.008237. This result is statistically significant at conventional levels (p < 0.05), leading us to reconsider our previous stance. With an F-test value of 2.1049 and degrees of freedom of 15 and 782 for the numerator and denominator, respectively, we find evidence that past values of news sentiment contain information that is useful in predicting future values of MSFT stock returns.

This finding calls for a reevaluation of the dynamic interplay between public sentiment, as captured by news, and subsequent investor behavior, as reflected in stock prices. It suggests that the informational content of news sentiment is indeed incorporated into market prices, albeit with a degree of complexity that may require further research to fully understand.

In light of this finding, our initial belief in the non-causal role of news sentiment may have been premature. This reinforces the need for a multifaceted approach to the analysis of financial time series, where causality may not always be apparent at first glance, but may emerge upon closer examination.
